### 2018-09-27 paper.tex
Wip

\lstset{
  basicstyle=\footnotesize\ttfamily,
  breaklines=true,
  captionpos=b,
  framextopmargin=5pt,
  framexbottommargin=5pt,
  frame=tb
}

@@

Figure \ref{fig:guild-file} illustrates a simple Guild file.

\begin{figure}
{\footnotesize
\begin{verbatim}
 model: mnist
 description:
   Logistic regression classifier
   of MNIST digits
 operations:
   train:
     main: train_mnist
     flags:
       batchs-size: 32
       train-steps: 1000
   evaluate:
     main: eval_mnist
\end{verbatim}}
\caption{Guild file example}
\label{fig:guild-file}
\end{figure}

@@

Figure \ref{fig:resource} illustrates a resource definition for a
model. The resource \verb|idx-data| defines four source files that
will be made available to any operation that requires that resource.

\begin{figure}
{\footnotesize
\begin{verbatim}
 model: mnist
 resources:
   idx-data:
     description:
       MNIST dataset in compressed
       IDX format
     sources:
       - train-images-idx3-ubyte.gz
       - train-labels-idx1-ubyte.gz
       - t10k-images-idx3-ubyte.gz
       - t10k-labels-idx1-ubyte.gz
\end{verbatim}}
\caption{Model resource example}
\label{fig:resource}
\end{figure}

@@

\begin{description}
    \item[user] person using a model developed by someone else
    \item[developer] person developing a model, either from scratch or
      adapting the work of others
    \item[packager] person neither using nor developing a model but
      who packages models from developers for use by users
\end{description}

\subsection{Publish a Model for Use}

A developer may encourage others to use her model by documenting the
steps needed to install, adapt, and use it. She may be motivated to
simplify the user experience so that her work is more easily and
succesfully applied. The benefit to her may be quality feedback from
expert users, improved reputation, a sense of contribution, or her
colleagues may simply expect that work be documented and easy to use.

@@

{\footnotesize
\begin{verbatim}
1. Install required packages
2. Download the raw data
3. Prepare the raw data for training
4. Train the model
5. Export the trained model for deployment
\end{verbatim}}

While these steps provide invaulable information, they still rely on
the user to carefully execute them and troubleshoot issues as they
arise. The process, while thoroughly documented, can take time and is
subject to error at any step.

Guild AI can be used to both document and automate these
steps. Listing~\ref{lst:published-model} illustrates a Guild file that
might be used to codify the steps as model operations, which are both
discoverable and runnable.

\begin{lstlisting}[caption=whatever,label={lst:published-model}]
model: sample
operations:
  prepare-data:
    description:
      Prepare data for training
    main: prepare_data
    requires: raw-data
  train:
    description: Train model
    main: train
    requires: prepared-data
  export:
    description:
      Export model for deployment
    main: export
    requires: trained-model
resources:
  raw-data:
    description: Raw data for training
    sources:
      - url: http://mydata.com/raw.tgz
  prepare-data:
    description: Prepared data
    sources:
      - operation: prepare-data
  trained-model:
    description: Trained model
    sources:
      - operation: train
\end{lstlisting}

@@

@@

@@

@@

>>>>

\lstset{
  aboveskip=12pt,
  basicstyle=\footnotesize\ttfamily,
  breaklines=true,
  captionpos=b,
  frame=tb,
  framexbottommargin=5pt,
  framextopmargin=5pt,
}

@@

Listing~\ref{lst:guild-file} illustrates a simple Guild file.

\begin{lstlisting}[
    caption=Simple Guild file example,
    label={lst:guild-file}]
model: mnist
description:
  Logistic regression classifier
  of MNIST digits
operations:
  train:
    main: train_mnist
    flags:
      batchs-size: 32
      train-steps: 1000
  evaluate:
    main: eval_mnist
\end{lstlisting}

@@

Listing~\ref{lst:resource} illustrates a resource definition for a
model. The resource \verb|idx-data| defines four source files that
will be made available to any operation that requires that resource.

\begin{lstlisting}[
    caption=Model resource example,
    label={lst:resource}]
model: mnist
resources:
  idx-data:
    description:
      MNIST dataset in compressed
      IDX format
    sources:
      - train-images-idx3-ubyte.gz
      - train-labels-idx1-ubyte.gz
      - t10k-images-idx3-ubyte.gz
      - t10k-labels-idx1-ubyte.gz
\end{lstlisting}

@@

\begin{description}
    \item[user] person using a model developed by someone else
    \item[developer] person developing a model, either from scratch or
      adapting the work of others
    \item[packager] person neither using nor developing a model but
      who packages models created by developers for use by users
\end{description}

\subsection{Publish a Model for Use}
\label{sec:publish-model}

A \emph{developer} may encourage others to use her model by
documenting the steps needed to install, adapt, and use it. She may be
motivated to simplify the user experience so that her work is more
easily and succesfully applied. The benefit to her may be quality
feedback from expert users, improved reputation, a sense of
contribution, or her colleagues may simply expect that work be
documented and easy to use.

@@

\begin{enumerate}
   \item Install required packages
   \item Download the raw data
   \item Prepare the raw data for training
   \item Train the model
   \item Export the trained model for deployment
\end{enumerate}

While these steps provide invaulable information, they still rely on
the user to carefully execute them and troubleshoot issues as they
arise. The process takes time and is subject to error at any step.

Guild AI can be used to both document and automate these
steps. Listing~\ref{lst:published-model} illustrates a Guild file that
might be used to codify the steps as model operations.

\begin{lstlisting}[
    caption=Sample project Guild file,
    label={lst:published-model}]
model: sample
operations:
  prepare-data:
    description:
      Prepare data for training
    main: prepare_data
    requires: raw-data
  train:
    description: Train model
    main: train
    requires: prepared-data
  export-model:
    description:
      Export model for deployment
    main: export_model
    requires: trained-model
resources:
  raw-data:
    description: Raw data for training
    sources:
      - url: http://mydata.com/raw.tgz
  prepare-data:
    description: Prepared data
    sources:
      - operation: prepare-data
  trained-model:
    description: Trained model
    sources:
      - operation: train
\end{lstlisting}

The operations can be run using the Guild AI command line interface as
follows:

{\footnotesize
\begin{verbatim}
 $ guild run prepare-data
 $ guild run train
 $ guild run export-model
\end{verbatim}}

Furthermore, Guild AI provides model help, detailing available
operations, supported flags, and operation dependencies.

A user can get help for the model by running:

{\footnotesize
\begin{verbatim}
 $ guild help
\end{verbatim}}

Documentation is automatically generated from the model description,
making Guild AI models self documenting.

The developer can chose to distribute her model as a
project---e.g. making it available as a GitHub repository. She may
alternatively package the model and deploy as an installable Python
wheel distribution artifact.

@@

The flip side of the use case outline in
Section~\ref{sec:publish-model} is that of the \emph{user}. A user is
interesting in taking the work of a model developer and applying it,
either by adapating her model to a new application or by integrating
it as into another project.

Users commonly rely on thorough documentation to effectively use
models. If a project is under documented or otherwise hard to work
with, it may not be used. This is a lost opportunity for both user and
developer that hinders adoption and model improvement.

As described in Section~\ref{sec:publish-model}, a published model is
simple to use. From the standpoint of the user, the model can be used
either by cloning the project repository---e.g. as in the case of
GitHub published models---or install the model as a Python
package. Assuming the project was published as a package named
\verb|sample-project|, the user might run these commands to use the
model:

\setlength{\parindent}{1em}

{\footnotesize\emph{Install the package}}

{\footnotesize
\begin{verbatim}
 $ guild install sample-project
\end{verbatim}}

{\footnotesize\emph{View package help}}

{\footnotesize
\begin{verbatim}
 $ guild help sample-package
\end{verbatim}}

{\footnotesize\emph{Use the model}}

{\footnotesize
\begin{verbatim}
 $ guild run sample:prepare-data
 $ guild run sample:train
 $ guild run sample:export-model
\end{verbatim}}

\setlength{\parindent}{0em}

This illustrates the usefulness of packaging models for
reuse. Packaged models in this case consist of more than the
underlying model code---they include additional information that makes
them trivial to use relative to traditional forms of model reuse.

@@

Complex software ecosystems such as GNU/Linux have formalized the role
of a software \emph{packager} (or \emph{package maintainer}). The goal
of a packager is to improve the value of a software distribution by
creating high quality software packages that users find
useful. Packers are often independent of the upstream software
projects they package, serving as intermediaries betweeen the software
developers and distribution users.

Guild AI supports this pattern by allowing packagers to incorporate
upstream model sources into Guild packages, patching them as needed to
provide the desired functionality to users.

Consider a case where a model developer is overloaded with research
tasks and doesn't have time to provide documentation for using her
models or publishing them for use as described in
Section~\ref{sec:publish-model}. An independent \emph{packager} may
step up to create a package instead.

A packager must assume that upstream sources are available \emph{as
  is} and he cannot expect developers to fix bugs or add new features
(though he may certainly ask and hope). Packagers therefore must be
prepared to \emph{patch} model code as needed to get it to work as
expected.

Guild AI provides various means of patching model source to support
packagers in this role.

@@

The roles of \emph{developer}, \emph{user}, and \emph{packager} often
exist within a single organization or team. Guild AI supports team
collaboration by supporting each role as a separate concern.

Within a typical organization structure, the roles map as follows:

\setlength{\parindent}{1em}
developer $\rightarrow$ \emph{research scientist}

user $\rightarrow$ \emph{software engineer}

packager $\rightarrow$ \emph{research engineer}
\setlength{\parindent}{0em}

Of course this is a rough mapping and not all organizations share this
structure. Whatever the job titles and division of labor, Guild AI
supports the three facets of collaboration: model development, model
use, and intermediation between those two points in the interest of
freeing up developer time and resources to focus on pure model
development.

@@

In the case of a single developer who has no interest in publishing
models or otherwise supporting users, Guild AI model abstractions
provide value. By implementing models as units of operations, as is
the case with Guild AI model packaging, developers ensure their models
are easy to use from the early stages of development. This streamlines
the development process as developers can repeatedly run and iterate
through model operations using high level commands with well defined,
self documenting interfaces.
